---
title: "Drug and Health Plan Rating"
author: "Bowen Liu"
date: "March 3, 2016"
output:
  html_document: default
  pdf_document:
    number_sections: yes
  word_document: default
---
# Problem Description & Current State of Affairs
Medicare.gov released recent two years's data about "Drug and Health Plan" coverage around the country. The data inclues who, when, where, and how the plans were used. There are many plans with low ratings. And we want to find out how predictors such as political affiliation, income level, insurance rates, unemployment rates, hospital type, or happiness index contribute to the final rating of one specific type of plan. This is where we start to build the predicition modeling. It could be taken for reference for future plan upgrading or new plan release.

# Algorithms Used and Rationale
|  Algorithms  |                            Rationale                         |
|--------------|--------------------------------------------------------------|
|K-means, SOM| Before getting into the prediction modeling, we need first explore the structure of the datasets. K-means is one easy and light algorithms to finish this job. For further exploring, we might use SOM to do the same thing.|
|Multivariant Regession, Naive Bayes, Decision Tree, ANN|Since the project is one supervised problem, we are going to apply multiple prediction models to the datasets. And we will choose the final one with the highest accuracy.|

# Data Description: origin, data points, and vairables
|            Data Sets          |   Data Point | Variables                     |
|-------------------------------|--------------|-------------------------------|
|[PlanInfoCountyFipsCode.csv, vwStarRatingSummaryScores.csv](https://www.medicare.gov/download/downloaddb.asp) | 3,065,622 |"contract_id", "plan_id", "contract_year", "org_name", "plan_name", "city", "state_code", "zip_code", "CountyFIPSCode", "Summary_Score"|
|[Health Insurance statistics](http://www.census.gov/data/developers/data-sets/Health-Insurance-Statistics.html) | 3,143 |"NAME", "NIC_PT"(Number Insured), "NUI_PT(Number Uninsured)", "PCTIC_PT"(Percent Insured), "PCTUI_PT"(Percent Uninsured), "state", "county", "CountyFPISCode"|
|[Unemployment Rate](http://zipatlas.com/us/zip-code-comparison/unemployment-rate.htm) | 31,706 | "Zip Code", "Unemployment Rate" |
|[Income by zipcode](https://www.irs.gov/uac/SOI-Tax-Stats-Individual-Income-Tax-Statistics-2013-ZIP-Code-Data-(SOI)) |27,791 | "zip code", "AGI_STUB"(size of adjusted gross income)|
|[Hospital General Information](https://data.medicare.gov/data/hospital-compare) | 4,806 | "zipcode", "Hospital Ownership"(Government, non-profit, Proprietary, ) |

# Durg and Health Plan with copay, deductible, coinsurance
```{r,cache=TRUE,warning=FALSE}
# load vwPlanDrugsCostSharing.csv
drug_cost <- read.csv("./data/2015Med2000_flatfiles/vwPlanDrugsCostSharing.csv", stringsAsFactors = FALSE, na.strings = c("Information Not Available", ""))
# keep the copay, deductible, and coinsurance
colnames <- c("contract_id", "plan_id", "mnthly_prm", "ann_ddctbl", "copay_amt_inn_prefrd", "coins_amt_inn_prefrd")
drug_cost2 <- drug_cost[,colnames]
drug_cost3 <- drug_cost2
# split the range into standalone characters
drug_cost3$copay_low <- sapply(strsplit(drug_cost2$copay_amt_inn_prefrd, split="-"), '[', 1)
drug_cost3$copay_high <- sapply(strsplit(drug_cost2$copay_amt_inn_prefrd, split="-"), '[', 2)

drug_cost3$coin_low <- sapply(strsplit(drug_cost2$coins_amt_inn_prefrd, split="-"), '[', 1)
drug_cost3$coin_high <- sapply(strsplit(drug_cost2$coins_amt_inn_prefrd, split="-"), '[', 2)
# remove $ and %
library(gdata)
drug_cost3$copay_low <- gsub("^\\$", "", trim(drug_cost3$copay_low))
drug_cost3$copay_high <- gsub("^\\$", "", trim(drug_cost3$copay_high))
drug_cost3$coin_low <- gsub("%$", "", trim(drug_cost3$coin_low))
drug_cost3$coin_high <- gsub("%$", "", trim(drug_cost3$coin_high))
# change NA to 0
drug_cost3$copay_low[is.na(drug_cost3$copay_low)] <- 0
drug_cost3$copay_high[is.na(drug_cost3$copay_high)] <- 0
drug_cost3$coin_low[is.na(drug_cost3$coin_low)] <- 0
drug_cost3$coin_high[is.na(drug_cost3$coin_high)] <- 0
# choose the maxium value
drug_cost3$copay_max <- apply(drug_cost3[, c("copay_low", "copay_high")], 1, max)
drug_cost3$coin_max <- apply(drug_cost3[, c("coin_low", "coin_low")], 1, max)
# subset the needed matrix
drug_cost4 <- drug_cost3[, c("contract_id", "mnthly_prm", "ann_ddctbl", "copay_max", "coin_max")]
# change the column type to numeric
drug_cost4$ann_ddctbl <- as.numeric(drug_cost4$ann_ddctbl)
drug_cost4$copay_max <- as.numeric(drug_cost4$copay_max)
drug_cost4$coin_max <- as.numeric(drug_cost4$coin_max)
drug_cost4$mnthly_prm <- as.numeric(drug_cost4$mnthly_prm)
# aggreate the drug_cost by contract id
drug_cost4_aggr <- aggregate(drug_cost4, by = list(drug_cost4$contract_id), FUN = median)
drug_cost4_aggr <- drug_cost4_aggr[, c("Group.1", "mnthly_prm", "ann_ddctbl", "copay_max", "coin_max")]
colnames(drug_cost4_aggr) <- c("contract_id", "mnthly_prm", "ann_ddctbl", "copay_max", "coin_max")
#write.csv(drug_cost4_aggr[, c("Group.1", "copay_max", "coin_max")], file = "./drug_cost.csv")
str(drug_cost4_aggr)
head(drug_cost4_aggr)
```
# Drug and Health Cost Tier Levle
```{r,cache=TRUE,warning=FALSE}
# load vwPlanDrugTierCost.csv 
cost_tier <- read.csv("./data/2015Med2000_flatfiles/vwPlanDrugTierCost.csv", stringsAsFactors = FALSE)
# keep contract_id, plan_id, languange_id, tier_lable
colnames <- c("contract_id", "language_id", "tier_label")
cost_tier2 <- cost_tier[, colnames]
cost_tier2 <- cost_tier2[cost_tier2$language_id == "1", ]
cost_tier3 <- cost_tier2[, c("contract_id", "tier_label")]
# count unique tier numbers for each contract id
cost_tier4 <- aggregate(data = cost_tier3,  tier_label ~ contract_id, function(x) length(unique(x)))
#cost_tier5 <- aggregate(cost_tier4, by = list(cost_tier4$contract_id), FUN = sum)
str(cost_tier4)
head(cost_tier4)
```
# Durg and Health Plan Ratings
```{r,cache=TRUE}
options(width=80)
# load rating data
ratingdata <- read.csv("./data/2015StarRatings_flatfiles/vwStarRating_SummaryScores.csv",na.strings = c("Plan too new to be measured", "Not enough data available"), stringsAsFactors = FALSE)
# remove Spanish Version
ratingdata2 <- ratingdata[ratingdata$lang_dscrptn=="English",]
# keep the columns of  "Contract_ID", "Summary_Score", "Star_Rating_Current"
colnames <- c("Contract_ID", "Summary_Score", "Star_Rating_Current")
ratingdata2 <- ratingdata2[, colnames]
# split the string to get the rating number
ratingdata2$Star_Rating_Current <- sapply(strsplit(ratingdata2$Star_Rating_Current, split=" "), '[', 1)
#ratingdata2$Star_Rating_Previous <- sapply(strsplit(ratingdata2$Star_Rating_Previous, split=" "), '[', 1)
# remove NAs of the current 2015 year
ratingdata2_nocurrna <- ratingdata2[(!is.na(ratingdata2$Star_Rating_Current)), ]
str(ratingdata2_nocurrna)
head(ratingdata2_nocurrna)
# subset the rating data
# ratingdata3 <- ratingdata2[ratingdata2$Summary_Score=="Overall Plan Rating", 1:3]
#head(ratingdata3)

#join plan and ratings by Contract ID
# med_data <- merge(x=data, y=ratingdata3, by.x = "contract_id", by.y = "Contract_ID", all.x = TRUE)
# sum(is.na(med_data[med_data$Summary_Score]))
# nrow(med_data)
#tail(med_data, 10)
```
# Geographic data
```{r,cache=TRUE}
zip_contract <- read.csv("./data/2015Med2000_flatfiles/vwLocalContractServiceAreas.csv", stringsAsFactors = FALSE)
zip_contract2 <- zip_contract[, c("Zip_Code", "Contract_ID")]
```
# Internal: combine by contract id
```{r,cache=TRUE}
cost <- merge(cost_tier4, drug_cost4_aggr, all = TRUE, by = "contract_id")
cost_rating <- merge(cost, ratingdata2_nocurrna, all = TRUE, by.x = "contract_id", by.y = "Contract_ID")
cost_rating$tier_label[is.na(cost_rating$tier_label)] <- 0
cost_rating2 <- na.omit(cost_rating)
cost_rating_zip <- merge(cost_rating2, zip_contract2, all = TRUE, by.x = "contract_id", by.y = "Contract_ID")
cost_rating_zip2 <- na.omit(cost_rating_zip)
str(cost_rating_zip2)
#ratingdata2_nocurrna
#cost_tier4
#zip_contract2
#drug_cost4_aggr
```
# Unemployment Rate by Zipcode
```{r,cache=TRUE}
unemployment_data <- read.table("./data/unemployment.txt", stringsAsFactors = FALSE)
unemployment_data2 <- unemployment_data[, c(2, 6, 7)]
colnames(unemployment_data2) <- c("zipcode", "polulation", "unemployment_rate")
unemployment_data2$polulation <- as.numeric(gsub("[[:punct:]]", "", unemployment_data2$polulation))
unemployment_data2$unemployment_rate <- as.numeric(unemployment_data2$unemployment_rate)

str(unemployment_data2)
head(unemployment_data2)
```
# Total Income by Zip Code
```{r,cache=TRUE}
# loading income level, we only need State, zipcode, agi_stub(levels: 1,2,3,4,5,6)
# 1 = $1 under $25,000, 2 = $25,000 under $50,000, 3 = $50,000 under $75,000
# 4 = $75,000 under $100,000, 5 = $100,000 under $200,000, 6 = $200,000 or more
income_data <- read.csv("./data/incomelevel_zipcode2013/zipcodenoagi13.csv", stringsAsFactors = FALSE)
income_data2 <- income_data[, c("STATE", "ZIPCODE", "A02650")]
colnames(income_data2)[3] <- "total_income"
str(income_data2)
head(income_data2)
```
# External: combine data by zip code or county
```{r,cache=TRUE}
unemploy_Income <- merge(unemployment_data2, income_data2, all = TRUE, by.x = "zipcode", by.y = "ZIPCODE")
unemploy_Income_0na <- na.omit(unemploy_Income)
unemploy_Income_0na$averge_income <- unemploy_Income_0na$total_income / unemploy_Income_0na$polulation
#unemploy_Income_0na <- unemploy_Income[!is.na(unemploy_Income$unemployment_rate), ]
#str(unemploy_Income_0na)
#head(unemploy_Income_0na)
#write.csv(unemploy_Income_0na, file = "./external_data.csv")
```
# Combine all data
```{r,cache=TRUE}
final_data <- merge(unemploy_Income_0na, cost_rating_zip2, all = TRUE, by.x = "zipcode", by.y = "Zip_Code")
final_data <- na.omit(final_data)
# unemploy_Income_0na
# cost_rating_zip2
```
# Test the model
```{r fig.width=7, fig.height=6,cache=TRUE}
colnames(final_data)
library(C50)
# split into train and test sets
sample_size <- floor(0.8 * nrow(final_data))
training_index <- sample(seq_len(nrow(final_data)), size = sample_size)

train <- final_data[training_index,]
test <- final_data[-training_index,]
predictors <- c('unemployment_rate', 'averge_income', 'tier_label', 'mnthly_prm', 'ann_ddctbl', 'copay_max', 'coin_max')
train$Star_Rating_Current <- as.factor(train$Star_Rating_Current)
model <- C5.0.default(x = train[,predictors], y = train$Star_Rating_Current)
plot(model)

pred <- predict(model, newdata = test)
evaluation <- cbind(test, pred)
ncol(evaluation)
evaluation$correct <- ifelse(evaluation$Star_Rating_Current == evaluation$pred,1,0)
sum(evaluation$correct)/nrow(evaluation)
```