---
title: "Drug and Health Plan Rating"
author: "Bowen Liu"
date: "March 3, 2016"
output:
  html_document: default
  pdf_document:
    number_sections: yes
  word_document: default
---
# Problem Description & Current State of Affairs
Medicare.gov released recent two years's data about "Drug and Health Plan" coverage around the country. The data inclues who, when, where, and how the plans were used. There are many plans with low ratings. And we want to find out how predictors such as political affiliation, income level, insurance rates, unemployment rates, hospital type, or happiness index contribute to the final rating of one specific type of plan. This is where we start to build the predicition modeling. It could be taken for reference for future plan upgrading or new plan release.

# Algorithms Used and Rationale
|  Algorithms  |                            Rationale                         |
|--------------|--------------------------------------------------------------|
|K-means, SOM| Before getting into the prediction modeling, we need first explore the structure of the datasets. K-means is one easy and light algorithms to finish this job. For further exploring, we might use SOM to do the same thing.|
|Multivariant Regession, Naive Bayes, Decision Tree, ANN|Since the project is one supervised problem, we are going to apply multiple prediction models to the datasets. And we will choose the final one with the highest accuracy.|

# Data Description: origin, data points, and vairables
|            Data Sets          |   Data Point | Variables                     |
|-------------------------------|--------------|-------------------------------|
|[PlanInfoCountyFipsCode.csv, vwStarRatingSummaryScores.csv](https://www.medicare.gov/download/downloaddb.asp) | 3,065,622 |"contract_id", "plan_id", "contract_year", "org_name", "plan_name", "city", "state_code", "zip_code", "CountyFIPSCode", "Summary_Score"|
|[Health Insurance statistics](http://www.census.gov/data/developers/data-sets/Health-Insurance-Statistics.html) | 3,143 |"NAME", "NIC_PT"(Number Insured), "NUI_PT(Number Uninsured)", "PCTIC_PT"(Percent Insured), "PCTUI_PT"(Percent Uninsured), "state", "county", "CountyFPISCode"|
|[Unemployment Rate](http://zipatlas.com/us/zip-code-comparison/unemployment-rate.htm) | 31,706 | "Zip Code", "Unemployment Rate" |
|[Hospital General Information](https://data.medicare.gov/data/hospital-compare) | 4,806 | "zipcode", "Hospital Ownership"(Government, non-profit, Proprietary, ) |
|[Income by zipcode](https://www.irs.gov/uac/SOI-Tax-Stats-Individual-Income-Tax-Statistics-2013-ZIP-Code-Data-(SOI)) |27,791 | "zip code", "AGI_STUB"(size of adjusted gross income)|

# Data Preprocessing of Drug and Health Plan, Insurance, Income Level, and Hospital Type
Prepare Durg and Health Plan By County and State
```{r,cache=TRUE}
# load PlanInfoCounty_FipsCodeLessThan30000
data1 <- read.csv("./data/2016Med2000_PlanInfoCounty1_flatfiles/PlanInfoCounty_FipsCodeLessThan30000.csv", stringsAsFactors = FALSE)
colnames <- c("contract_id", "plan_id", "contract_year", "org_name", "plan_name", "city", "state_code", "zip_code", "CountyFIPSCode")
data12 <- data1[,colnames]
# load PlanInfoCounty_FipsCodeMoreThan30000
data2 <- read.csv("./data/2016Med2000_PlanInfoCounty2_flatfiles/PlanInfoCounty_FipsCodeMoreThan30000.csv", stringsAsFactors = FALSE)
data22 <- data2[,colnames]
# join 2 parts of less than 30000 and more than 30000
data <- rbind(data12, data22)
# write up plan info by county_FIPSCode
#write.csv(data, file = "./PlanInfoCounty_FipsCode.csv")
#head(data)
```
Prepare Durg and Health Plan Ratings
```{r,cache=TRUE}
options(width=80)
# load rating data
ratingdata <- read.csv("./data/2016StarRatings_flatfiles/vwStarRating_SummaryScores.csv",na.strings = c("Plan too new to be measured", "Not enough data available"), stringsAsFactors = FALSE)
# remove Spanish Version
ratingdata2 <- ratingdata[ratingdata$lang_dscrptn=="English",]
# split the string to get the rating number
ratingdata2$Star_Rating_Current <- sapply(strsplit(ratingdata2$Star_Rating_Current, split=" "), '[', 1)
ratingdata2$Star_Rating_Previous <- sapply(strsplit(ratingdata2$Star_Rating_Previous, split=" "), '[', 1)
# subset the rating data
ratingdata3 <- ratingdata2[ratingdata2$Summary_Score=="Overall Plan Rating", 1:3]
#head(ratingdata3)

#join plan and ratings by Contract ID
# med_data <- merge(x=data, y=ratingdata3, by.x = "contract_id", by.y = "Contract_ID", all.x = TRUE)
# sum(is.na(med_data[med_data$Summary_Score]))
# nrow(med_data)
#tail(med_data, 10)
```
Prepare Insurance Coverage Data
```{r,cache=TRUE}
#Use census.gov developer API to get the insurance related data
#http://api.census.gov/data/timeseries/healthins/sahie?get=NAME,NIC_PT,NUI_PT,PCTIC_PT,PCTUI_PT&for=COUNTY:*&time=2013
#install.packages("rjson")
library(rjson)
json_file <- "./sahie.json"
json_data <- data.frame(fromJSON(file = json_file))
insurance_data <- t(json_data)
rownames(insurance_data) <- NULL

colnames(insurance_data) <- insurance_data[1,]
insurance_data <- data.frame(insurance_data[-1,], stringsAsFactors = FALSE)

insurance_data$CountyFPISCode <- as.numeric(with(insurance_data, paste(state, county, sep = "")))
#write.csv(json_data1, file = "./insurance_coverage_2013.csv")
insurance_data2 <- insurance_data[, c("NAME", "NIC_PT", "PCTIC_PT", "CountyFPISCode")]
#head(insurance_data2)
```
Prepare Income Level by Zip Code
```{r,cache=TRUE}
# loading income level, we only need State, zipcode, agi_stub(levels: 1,2,3,4,5,6)
# 1 = $1 under $25,000, 2 = $25,000 under $50,000, 3 = $50,000 under $75,000
# 4 = $75,000 under $100,000, 5 = $100,000 under $200,000, 6 = $200,000 or more
income_data <- read.csv("./data/incomelevel_zipcode2013/13zpallagi.csv", stringsAsFactors = FALSE)
income_data2 <- income_data[, c("STATE", "zipcode", "agi_stub")]
#head(income_data2)
```
Prepare Hospital Type by Zip Code
```{r,cache=TRUE}
# loading hospital data, we only need 
hospital_data <- read.csv("./data/Hospital_Revised_Flatfiles/Hospital General Information.csv", stringsAsFactors = FALSE)
hospital_data2 <- hospital_data[, c("Hospital.Name", "State", "ZIP.Code", "County.Name", "Hospital.Type")]
#head(hospital_data2)
```
# Combine data
All the data sets would be combined by the same field of zip code.